% file: xxxx.tex (tpl-41)
% -
% Copyright - Glen Ritchie 2025
% License terms: MIT Free for template
%\documentclass[14pt]{extarticle}
\documentclass[12pt]{article}
\usepackage[small,sf,bf]{titlesec}%0pt
\usepackage{amsmath,amssymb,amsfonts,latexsym}
\usepackage[tmargin=1.0in,bmargin=1.0in,lmargin=1.0in,rmargin=1.0in]{geometry}
% Going with 'extarticle' and 14 point, for reading script purposes.

% ============================================================
\usepackage{setspace,graphics,rotating,amsthm}
\usepackage{lipsum,hyperref,hyperxmp,graphicx}
% csquotes conflicts with 'graphicx', if it is used?
% graphicx best for PDF import with XeLaTeX or EPS with latex only build program.
\usepackage{enumerate,xspace,calc,pifont,caption}

\usepackage{tikz,pgfplots,pgfplotstable}
\usepackage{fancyhdr}
\usepackage{fontspec}
\setmainfont{Times New Roman}

\newcommand*\aakeywa{%
Writing% <-- problem
}
\newcommand{\aatita}{The Pre-Pub Paper}%
\newcommand{\aaauta}{Glen Ritchie}%

\hypersetup{
    pdftitle    = {\aatita},
    pdfauthor   = {\aaauta},
    pdfsubject  = {Paper on Number Theory \& number triangles},
    pdfkeywords = {Number Theory, number triangles, math}
}

\author{\aaauta}
\title{\aatita}
\newcommand{\thesubtitle}{Undefined}

% ============================================================
\usepackage{cleveref}% Important - to be last one - called.
\usepackage{clever-ref}

% ============================================================
% (switch to other template!) \usepackage{euler-sym}
\usepackage{plain-sym}
\usepackage{writeB-macs}

\usepackage{pdfpages}

%\let\defpgsection\section
%\renewcommand{\section}{\newpage\defpgsection}%
% If want each section, on it's own page.

\newcommand{\aadots}{\dots\,}%
\newcommand{\aabart}{---\,\,}%

% Try to avoid sfrac from xfrac package.
\newcommand{\slfrac}[2]{\,^{#1}\!/_{#2}}

\newcommand{\aacenterrule}{%
  \begin{center}
    \rule{1in}{0.4pt}%
  \end{center}
}

% \setstretch{0.95}
% \setstretch{1.00}
%\doublespacing
\setlength{\parindent}{0pt}
\setlength{\parskip}{6.0pt}

\pagestyle{fancy}
\fancyhf{}              % clear all header and footer fields
%\fancyhead[C]{\thepage} % page number in upper right
\renewcommand{\headrulewidth}{0pt} % optional: remove header line
\renewcommand{\footrulewidth}{0pt} % optional: remove footer line
\fancyfoot[C]{\thepage}

% ============================================================

\begin{document}
 %\begin{abstract}
 %
 % I don't have an abstract. This is just my paper and there you go.
 %
 %\end{abstract}
 %\pagestyle{empty}
 
 \pagenumbering{gobble}

 \maketitle   
 
 %\tableofcontents
 \newpage
 \pagenumbering{arabic}
 
    \section{Presentation}
    \label{book:01}
 
     % ------------------------------------------------------------------
 
    \subsection{Preface}
    \label{sec:10}
 
     My name is Glen Ritchie. I have been a lifelong learner and student of Mathematics as a 
     hobby.
     
     I graduated from NDSU (2002) with a Bachelor of Science in Computer Science.
     
     I am writing this to explain the many number triangles that I have found. Plus, there is a 
     Number Triangle application written in Lazarus, a Pascal-derived language.
     
     For more on the NTA2 application. See it's section.
     
     So why study number triangles? I've been playing with Number Theory for decades. It's been 
     one of my primary interests. For instance, I wrote a paper on Pythagorean triples, which is 
     available online.
     
     Yet, my work on number triangles is novel.
     
     Many people have enabled me to undertake this work, and I give glory to God for it. I thank 
     my Lord Jesus.
     
     Glen Ritchie
     
     Bismarck, North Dakota - 2025

    \aacenterrule
 
     % ------------------------------------------------------------------
 
    \subsection{Introduction}
    \label{sec:11}

    The subject of my presentation is number triangles.

    Why study number triangles? The Binomial numbers and their patterns in Pascal's Triangle are 
    the answer. (Or Binomial coefficients.)
    
    Its equation is: (fill-in)
    
    Pascal's Triangle's addition rule with Binomial numbers is:
    
    (fill-in)
    
    How do you compute the Binomial numbers? Answer: It involves Factorials.
    
    (fill-in)
    
    I will not be studying Pascal's Triangle in depth, as Don Knuth notes in the next section. 
    There are a couple of thousand known patterns in binomial numbers and Pascal's Triangle.
    
    I will be studying the field of number triangles.
    
    I may need two meetings, to share most everything with you?
    
    I have tools ranging from an explanation of how created, then a diagram, then a Windows 
    application, plus an online dataset available to you.

    % ------------------------------------------------------------------
     
    \subsection{Don Knuth Quote}
    \label{sec:12}

    A quote on Pascal's Triangle and Binomial numbers --- discoveries.
    
    This is from page 54 of the 3rd Edition (most recent, as of 2025) TAOCP, Vol. Fundamental 
    Algorithms.
    
    \begin{quote}
    \textit{Binomial coefficients satisfy literally thousands of identities, and for centuries 
    their amazing properties have been continually explored.}
    
    \textit{In fact, there are so many relations present that when someone finds a new identity, 
    not many people get excited about it any more, except for the discoverer.}
    \end{quote}
    
    Persuade you that there is still interesting material in the subject. There are many number 
    triangles, with different patterns and types.
    
    Even though Don Knuth's quote is discouraging. I understand; mathematics is vast. It is 
    unlikely to find anything new. Yet, I have done this.
    
    I aim to investigate patterns and structural symmetries in number triangles and determine 
    whether these patterns can be connected to integer sequences in the OEIS database.
 
     % ------------------------------------------------------------------
     
    \subsection{Pascal's Triangle}
    \label{sec:13}

    Pascal's Triangle has been realized, found, discovered, and rediscovered several times. It 
    was known in China and the Islamic world before Pascal studied its properties and published 
    it.
    
    Blaise Pascal collected results from several mathematicians who had discovered it. Then, he 
    added more findings and realized their connection to Probability and Combinatorics.
    
        \subsubsection{RSD}    
               
        \begin{itemize}
            \item \textit{Including Rising Shallow Diagonal Diagram (RSD)}
        \end{itemize}
        
        What are they? Why is it important? How to calculate them?
        
        I have a diagram, which shows how they are arranged. It starts on the left-hand side of a 
        number triangle and slopes upwards.
        
        Is there a Binomial formula? Showing the calculation of an RSD sum in Pascal's Triangle?
        
        (fill-in)
        
        What are the applications of RSD values? Starting with it being an actual integer 
        sequence.
        
        \begin{enumerate}
            \item They are building blocks for infinite series. You can derive an approximate 
            series from an exact integer sequence. Then, apply it to a numerical calculation 
            algorithm.
            
            \item Data structures and runtime analyses (big-O notation) in Computer science, 
            often produce connections to integer sequences.
            
            \item Numbers surround us. Finance, Construction patterns, or scheduling routines all 
            involve simple integer patterns.
        \end{enumerate}
        
        \subsubsection{Number Triangle Format}
        
        \begin{itemize}
            \item \textit{Difference of Centered \& Left-Justified Number Triangles}
        \end{itemize}
        
        The following diagram, shows the difference between a centered number triangle and a 
        left-justified number triangle.
        
        I find the left-justified number triangle easier to read, in several ways. However, all 
        my number triangle diagrams are centered.
        
        \subsubsection{Simplex Numbers}
        
        \begin{itemize}
            \item \textit{Simplex Numbers \& Simplices}
        \end{itemize}
        
        What are the Simplex numbers? Refers to a family of ``Figurate numbers'' when 
        generalizing the polygonal numbers.
        
        \begin{enumerate}
            \item Natural numbers
            \item Triangular numbers
            \item Tetrahedral numbers
            \item Pentalope numbers
            \item and more.
        \end{enumerate}
        
        Are there patterns to them, in their formulas? Yes.
        
        (fill-in)
        
        Gradually, it becomes x times, (x+1) times, (x+2) times, etc.
        
        Additional division by $ 1 $, $ 1 $ times $ 2 $, times $ 3 $, times $ 4 $, ..
        
        They are very special numbers in Pascal's Triangle. However, we don't see them often, 
        beyond the Pentalope simplex numbers.
        
        \subsubsection{Geometric Relations}
        
        \begin{itemize}
            \item 0D: point.
            \item 1D: line segment.
            \item 2D: triangle.
            \item 3D: tetrahedron
            \item 4D: called 4-simplex, or 5-cell pentachoron, or pentatope.
        \end{itemize}

    % ------------------------------------------------------------------
     
    \subsection{Twice-Filled Pascal's Triangle}
    \label{sec:14}

    \begin{itemize}
        \item \textit{Twice-Filled Pascal's Triangle}
    \end{itemize}
    
    What if? I would ask you to construct a triangular number array, with Fibonacci numbers as 
    row sums and powers of $ 2 $ as the rising shallow diagonals?
    
    Does it exist? What would it prove? What does it look like?
    
    \begin{itemize}
        \item See the diagram.
    \end{itemize}
    
    First, it only calculates the odd Fibonacci numbers, not all of them.
    
    Second, it would imply a different kind of power series produces values of Fibonacci numbers, 
    not powers of $ 2 $.

        \subsubsection{OEIS Presence}
        
        The number triangle pattern is listed in OEIS. It is a sequence https://oeis.org/A046854. 
        Described as follows:
        
        \begin{itemize}
            \item Triangle read by rows: $ T(n, k) = binomial( floor( (n+k) / 2), k), 
            n >= k >= 0 $.
        \end{itemize}
        
        Is this a special number triangle? Perhaps, not. However, I find it very curious. It does 
        exist on OEIS, and by the nature of symmetry with Pascal's Triangle. It flips the RSD and 
        RS values, for the most part.

    % ------------------------------------------------------------------
   
    \subsection{Rudimentary Topics}
    \label{sec:15}

        \subsubsection{OEIS}
        
        The On-line Encyclopedia of Integer Sequences
        
        It is a free database of integer sequences stored to reveal. What integer sequence 
        connects to it?
        
        How is it useful? It allows anyone to identify integer sequences without knowing their 
        origins.
        
        \subsubsection{GIT \& GitHub}
        
        GIT and GitHub are two separate products used for revision control and for organizing 
        data repositories on GitHub.
        
        GIT is a tool for tracking changes in code or data and enabling collaboration.
        
        There are several freely available GIT GUI applications. SmartGit, in particular, allows 
        unrestricted non-commercial use of its GIT client software.
        
        GitHub is a website where GIT repositories are stored. The website fosters collaboration, 
        with several features.
        
        I encourage everyone to use these free tools and apply them on their computers. GitHub 
        even allows ``private'' repositories to free users.
        
        \subsubsection{\LaTeX{}}
        
        \LaTeX{} is a typesetting system for document publication, with a focus on high-quality 
        mathematical output.
        
        There is so much that anyone can say here. However, I believe it is taught on this campus 
        by the Department of Engineering.
        
        If you are interested in an alternative? \textit{Typst.app} appears to be very popular.

    % ------------------------------------------------------------------
    
    \subsection{Teardrop Number Triangle}
    \label{sec:16}

    Why, the name Teardrop? Simple. The pattern of solving numbers tends to fall and then grow 
    steadily.
    
        \subsubsection{Blackboard Challenge}
        
        What is the basis of Teardrop? A simple question.
        
        \begin{itemize}
            \item Is there a number triangle? Where the row sums (RS) don't matter and the rising 
            shallow diagonal (RSD) equals the Triangular numbers? Find it.
        \end{itemize}
        
        \subsubsection{Teardrop}
        
        While the objective for Teardrop is simple. Instead of the Fibonacci numbers being the 
        solution to a Rising Shallow Diagonal? Is the goal to find the same result for Triangular 
        numbers?
        
        Is there such a triangle? Is it simple or complicated? Had no idea? Perhaps I was 
        ignorant?
        
        However, it isn't hard. And it exists. It's not listed on OEIS. From it, I discovered a 
        new way to construct several number triangles.
        
        Pause.
        
        \subsubsection{``On Board'' Question}
        
        Explain the "on board" idea for the Introduction, as a teacher would.
        
        Show, demonstrate, and talk about:
        
        \begin{itemize}
            \item Natural numbers
            \item Triangular numbers
            \item Pascal's Triangle: centered form.
            \item Pascal's Triangle: left-justified form.
            \item Then. Explain the challenge of Teardrop. Give someone a few minutes to guess.
            \item We can give them a hint with: \_ = $ 1 $
            \item \_ = $ 3 $
            \item \_ + \_ = $ 6 $
            \item \_ + \_ = $ 10 $, and so on.
        \end{itemize}
        
        \subsubsection{Reveal of Teardrop Triangle}
        
        Here, it is. Almost beautiful.
        
        \subsubsection{History}
        
        Is this number triangle known to OEIS? It is not.

    % ------------------------------------------------------------------
    
    \subsection{Golden Triangle}
    \label{sec:17}

    What is special about the Golden triangle?

    % ------------------------------------------------------------------
    
    \subsection{NTA2 Application}
    \label{sec:18}

    There are several features to the Number Triangle Application (version) 2.
    
    \begin{itemize}
        \item It allows you to create, store, retrieve, and make RSD / RS sums calculations on a 
        relatively large number of triangles. $ \bullet $ How many? A googol cubed times $ 1 $ 
        trillion. If each cell holds a positive value from 1 to $ 10,000 $?
        
        \item It is an open-source software application for Windows 10/11. Made with Lazarus, 
        which is also an OSS development suite, based on the Pascal computer language.
        
        \item It has some form-filling wizards. The most powerful is the one that takes an 
        integer sequence (18 terms) and produces a number triangle. Where the RSD sum equals the 
        input values. The RS sum is where the surprise happens, when there is a connection.
        
        \item Extensive clipboard support is provided for the RSD and RS values. This is to 
        assist you in searching OEIS's database for matches to the sums.
        
        \item It computes RSD Rising Shallow Diagonals, RS Row sums, and even FSD Falling Shallow 
        Diagonals.
        
        \item There is even an option to shade the RSD fields gray when some number triangle 
        elements are missing, indicating they are incorrect.
        
        \item There is extensive cursor moving among the elements within the main form. Tabbing 
        moves across, and the up/down cursor moves up/down.
        
        \item A visual cue is left when you change the value of an element. Although if you enter 
        an element, then leave it. There is also a color change, and this gives you a trial. 
        Showing where you have been, in the number triangle elements.
        
        \item The file format saves more than the cells in the elements. It also saves the RSD \& 
        RS values, as well as the odd/even RSD and RS values. $ \bullet $ How is this useful? It 
        allows ``text scanning'' of \verb+.tri+ dataset files to search for numbers using any 
        \verb+grep+-like program, with your Operating System.
    \end{itemize}

        \subsubsection{General Introduction}
        
        \begin{enumerate}
            \item Overview / Purpose
            \item Key Benefits
            \item System Requirements
            \begin{enumerate}
                \item Unfortunately, it is not made for small-resolution monitors. Scrolling bars 
                will appear along the window edges.
            \end{enumerate}
            
            \item User Interface Tour
            \item Common Tasks / Workflows
            \item Tips, Shortcuts, and Best Practices
            \begin{enumerate}
                \item The most common operation is Ctrl+F for the Zipper \& Seam Construction 
                Wizard.
            \end{enumerate}
            
            \item Troubleshooting and Support
        \end{enumerate}
        
        Other Items. (Already discussed?)
        
        \begin{enumerate}
            \item How to use it?
            \item What is it?
            \item Who is it for?
            \item What problem does it solve?
        \end{enumerate}
        
        A suggested descriptive title: A Triangular Array Visualizer. (TAV)

% ## Future Issues
%
% (Comment out this section in LaTeX.)
%
% 1. Target audience and user personas for this Windows app
% 2. Key features to highlight in the app description
% 3. Step-by-step user guide outline for the app
% 4. Common troubleshooting steps and FAQs to include
% 5. How to write clear installation instructions for Windows
% 6. How to create feature walkthrough screenshots and captions
% 7. Which metric to show in the app's benefits section
% 8. How to write a concise app elevator pitch
% 9. Best structure for a software user manual table of contents
% 10. How to write release notes and changelogs for the app

    % ------------------------------------------------------------------
     
    \subsection{Vocabulary}
    \label{sec:19}   

% ------------------------------------------------------------------

% ============================================================    
    
\end{document}



